---
title: "Skilled Nursing Facility Benchmark Report"
subtitle: "Financial, utilization, payer-mix and labor benchmarking from CMS cost-report data"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
params:
  target_ccn: "465095"
  data_prefix: "national"
---

```{r setup}
library(dplyr)
library(jsonlite)
library(scales)
library(ggplot2)
library(gt)
library(glue)
library(readr)

aesc <- htmltools::htmlEscape
prefix <- as.character(params$data_prefix)
target_ccn <- as.character(params$target_ccn)
path <- function(stem) file.path("data/dashboard", paste0(prefix, "_", stem, ".json"))
required <- c(path("facility_index"), path("metric_specs"), path("facility_benchmarks"), path("facility_trends"), path("facility_strategy_flags"))
missing <- required[!file.exists(required)]
if (length(missing)) stop("Missing report inputs: ", paste(missing, collapse = ", "))

facilities <- fromJSON(path("facility_index"), simplifyDataFrame = TRUE)
metrics <- fromJSON(path("metric_specs"), simplifyDataFrame = TRUE)
benchmarks <- fromJSON(path("facility_benchmarks"), simplifyDataFrame = TRUE)
trends <- fromJSON(path("facility_trends"), simplifyDataFrame = TRUE)
flags <- fromJSON(path("facility_strategy_flags"), simplifyDataFrame = TRUE)

target <- facilities |> filter(as.character(.data$provider_ccn) == target_ccn) |> slice(1)
if (!nrow(target)) stop("Target CCN not found in national report data: ", target_ccn)
b <- benchmarks |> filter(as.character(.data$provider_ccn) == target_ccn)
t <- trends |> filter(as.character(.data$provider_ccn) == target_ccn)
f <- flags |> filter(as.character(.data$provider_ccn) == target_ccn)

peer_lookup_path <- file.path("data/processed", paste0("snf_peer_lookup_v041_", prefix, ".csv"))
latest_valid_path <- file.path("data/processed", "snf_latest_valid_reports_v041.rds")
peer_lookup <- if (file.exists(peer_lookup_path)) suppressMessages(read_csv(peer_lookup_path, show_col_types = FALSE)) else tibble()
peer_row <- peer_lookup |> filter(as.character(.data$provider_ccn) == target_ccn) |> slice(1)
peer_members <- tibble()
if (nrow(peer_row) && file.exists(latest_valid_path) && "peer_ccns" %in% names(peer_row)) {
  peer_ids <- trimws(unlist(strsplit(as.character(peer_row$peer_ccns[[1]]), ",", fixed = TRUE)))
  latest_valid <- readRDS(latest_valid_path)
  peer_members <- latest_valid |>
    filter(as.character(.data$provider_ccn) %in% peer_ids) |>
    arrange(.data$state, .data$facility_name) |>
    transmute(
      CCN = as.character(.data$provider_ccn),
      Facility = .data$facility_name,
      City = .data$city,
      State = .data$state,
      Beds = .data$total_beds,
      `Urban / rural` = .data$rural_urban,
      `Ownership / control` = .data$type_of_control
    )
}

fmt <- function(x, type = "number") {
  if (!length(x) || is.na(x)) return("—")
  if (identical(type, "percent")) return(percent(as.numeric(x), accuracy = 0.1))
  if (identical(type, "dollar")) return(dollar(as.numeric(x), accuracy = 1))
  number(as.numeric(x), accuracy = 0.1, big.mark = ",")
}
safe_text <- function(x) if (!length(x) || is.na(x) || !nzchar(trimws(as.character(x)))) "—" else as.character(x)

br <- function(metric) b |> filter(.data$metric == metric) |> slice(1)
metric_text <- function(metric, col = "target_value") {
  z <- br(metric)
  if (!nrow(z) || !col %in% names(z) || is.na(z[[col]][[1]])) return("—")
  fmt(z[[col]][[1]], z$format[[1]])
}
metric_row <- function(metric) metrics |> filter(.data$metric == metric) |> slice(1)
metric_field <- function(metric, field, fallback = "") {
  z <- metric_row(metric)
  if (!nrow(z) || !field %in% names(z) || is.na(z[[field]][[1]]) || !nzchar(as.character(z[[field]][[1]]))) return(fallback)
  as.character(z[[field]][[1]])
}
metric_explanation_html <- function(metric) {
  definition <- metric_field(metric, "definition", "This metric is calculated from the facility's CMS cost report.")
  strategic <- metric_field(metric, "strategic_use", "Use it with the peer comparison and longitudinal trend to identify questions for further review.")
  trend_note <- metric_field(metric, "trend_interpretation", "Focus on sustained changes rather than a single-year movement.")
  caveat <- metric_field(metric, "caveats", "Interpret this measure with other operating and financial context.")
  glue("<div class='metric-explain'><div><strong>What it measures:</strong> {aesc(definition)}</div><div><strong>Why it matters:</strong> {aesc(strategic)}</div><div><strong>How to read the trend:</strong> {aesc(trend_note)}</div><div class='metric-caveat'><strong>Keep in mind:</strong> {aesc(caveat)}</div></div>")
}

peer_definition <- if (nrow(b) && "peer_definition" %in% names(b)) b$peer_definition[[1]] else "Matched peers"
peer_count <- if (nrow(b) && "peer_count" %in% names(b)) b$peer_count[[1]] else NA_integer_
peer_mode <- if (nrow(b) && "peer_mode" %in% names(b)) b$peer_mode[[1]] else if (nrow(peer_row) && "peer_mode" %in% names(peer_row)) peer_row$peer_mode[[1]] else "auto"
peer_mode_label <- dplyr::case_when(
  peer_mode == "explicit" ~ "User-selected facilities",
  peer_mode == "filters" ~ "User-defined filters",
  TRUE ~ "Automatic matched peers"
)
latest_year <- if (nrow(t) && any(is.finite(suppressWarnings(as.numeric(t$source_year))))) suppressWarnings(max(as.numeric(t$source_year), na.rm = TRUE)) else target$source_year[[1]]

plot_metric <- function(metric, title) {
  spec <- metrics |> filter(.data$metric == metric) |> slice(1)
  d <- t |> filter(.data$metric == metric) |> arrange(.data$source_year) |>
    mutate(target_plot = ifelse(dplyr::coalesce(.data$target_has_report, TRUE), .data$target_value, NA_real_))
  if (!nrow(d)) return(NULL)

  p <- ggplot(d, aes(x = .data$source_year)) +
    geom_ribbon(aes(ymin = .data$peer_p25, ymax = .data$peer_p75, fill = "Peer IQR"), alpha = 0.38, color = NA) +
    geom_line(aes(y = .data$peer_median, color = "Peer median"), linetype = "22", linewidth = 0.9, na.rm = TRUE) +
    geom_point(aes(y = .data$peer_median, color = "Peer median"), shape = 15, size = 2.2, na.rm = TRUE) +
    geom_line(aes(y = .data$target_plot, color = "Target facility"), linewidth = 1.2, na.rm = TRUE) +
    geom_point(aes(y = .data$target_plot, color = "Target facility"), shape = 16, size = 2.8, na.rm = TRUE) +
    scale_color_manual(NULL, values = c("Target facility" = "#155EEF", "Peer median" = "#D97706")) +
    scale_fill_manual(NULL, values = c("Peer IQR" = "#CBD5E1")) +
    guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
    labs(title = title, subtitle = "Target facility vs peer median; shaded band = peer 25th–75th percentile", x = NULL, y = NULL) +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 7)) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      legend.position = "top",
      legend.justification = "left",
      legend.box.spacing = grid::unit(0, "pt")
    )
  if (nrow(spec) && spec$format[[1]] == "percent") p <- p + scale_y_continuous(labels = percent_format(accuracy = 1))
  if (nrow(spec) && spec$format[[1]] == "dollar") p <- p + scale_y_continuous(labels = dollar_format(accuracy = 1))
  p
}
```

<style>
.report-hero {border:1px solid #d9e1ea;border-radius:12px;padding:1rem 1.2rem;margin-bottom:1.2rem;background:#f8fafc}
.report-kicker {color:#667085;font-size:.9rem;margin-bottom:.35rem}.report-title {font-size:1.35rem;font-weight:750}.report-meta {color:#475467;margin-top:.3rem}
.metric-grid {display:grid;grid-template-columns:repeat(auto-fit,minmax(155px,1fr));gap:.7rem;margin:1rem 0 1.4rem}.metric-card {border:1px solid #e1e6ed;border-radius:10px;padding:.8rem;background:white}.metric-label {font-size:.78rem;color:#667085}.metric-value {font-size:1.35rem;font-weight:750;margin-top:.15rem}
.profile-grid {display:grid;grid-template-columns:repeat(auto-fit,minmax(180px,1fr));gap:.65rem;margin:.8rem 0 1.2rem}.profile-item {border-bottom:1px solid #eaecf0;padding:.45rem .15rem}.profile-label {font-size:.76rem;color:#667085;text-transform:uppercase;letter-spacing:.03em}.profile-value {font-weight:650;margin-top:.15rem}
.flag {border-left:5px solid #98a2b3;padding:.65rem .85rem;margin:.6rem 0;background:#fafafa;border-radius:6px}.flag.Risk{border-left-color:#b42318}.flag.Watch{border-left-color:#dc6803}.flag.Strength{border-left-color:#027a48}
.peer-box {border:1px solid #d9e1ea;border-radius:10px;padding:.9rem 1rem;background:#f8fafc;margin:.8rem 0 1.2rem}.peer-warning {border-left:4px solid #dc6803;padding:.55rem .75rem;background:#fffaeb;margin:.6rem 0}
.metric-explain {border-left:4px solid #98a2b3;background:#f8fafc;padding:.75rem .9rem;margin:.25rem 0 1.5rem;font-size:.93rem;line-height:1.45}.metric-explain > div {margin:.2rem 0}.metric-caveat {color:#475467}
.source-note{font-size:.84rem;color:#667085;border-top:1px solid #e1e6ed;padding-top:1rem;margin-top:2rem}
</style>

```{r results='asis'}
cat(glue("<div class='report-hero'><div class='report-kicker'>SNFBenchmarking.com</div><div class='report-title'>{aesc(target$facility_name[[1]])}</div><div class='report-meta'>CCN {aesc(target_ccn)} · {aesc(target$city[[1]])}, {aesc(target$state[[1]])} · Latest report year {latest_year}</div><div class='report-meta'>Peer method: {aesc(peer_mode_label)} · {aesc(peer_definition)} · n = {peer_count}</div></div>"))

cards <- list(
  c("Occupancy", metric_text("occupancy_rate")),
  c("Medicare day share", metric_text("medicare_day_share")),
  c("Revenue / resident day", metric_text("net_patient_revenue_per_day")),
  c("Operating expense / day", metric_text("operating_expense_per_day")),
  c("Patient service margin", metric_text("patient_service_margin")),
  c("Contract labor ratio", metric_text("contract_labor_ratio")),
  c("Labor cost / day", metric_text("total_labor_cost_per_day")),
  c("Days cash on hand", metric_text("days_cash_on_hand"))
)
cat("<div class='metric-grid'>")
for (x in cards) cat(glue("<div class='metric-card'><div class='metric-label'>{aesc(x[[1]])}</div><div class='metric-value'>{aesc(x[[2]])}</div></div>"))
cat("</div>")
```

# Facility profile

```{r results='asis'}
profile <- list(
  c("Licensed / reported beds", fmt(target$total_beds[[1]], "number")),
  c("Bed-size band", safe_text(target$bed_size_band[[1]])),
  c("Urban / rural", safe_text(target$rural_urban[[1]])),
  c("Ownership / control", safe_text(target$type_of_control[[1]])),
  c("Average daily census", metric_text("average_daily_census")),
  c("Report days", fmt(target$report_days[[1]], "number")),
  c("Report-period status", safe_text(target$report_period_status[[1]])),
  c("SNF / NF mix", safe_text(target$snf_nf_mix[[1]]))
)
cat("<div class='profile-grid'>")
for (x in profile) cat(glue("<div class='profile-item'><div class='profile-label'>{aesc(x[[1]])}</div><div class='profile-value'>{aesc(x[[2]])}</div></div>"))
cat("</div>")
if (!is.na(target$data_quality_flags[[1]]) && nzchar(as.character(target$data_quality_flags[[1]]))) {
  cat(glue("<div class='peer-warning'><strong>Data-quality note:</strong> {aesc(as.character(target$data_quality_flags[[1]]))}</div>"))
}
```

# Strategic signals

These signals are screening indicators generated from the facility's latest cost report and peer benchmark. They identify questions worth investigating; they are not guaranteed savings estimates or operational conclusions.

```{r results='asis'}
if (!nrow(f)) {
  cat("<div class='flag'>No rule-based strategic flags were generated for this report.</div>")
} else {
  for (i in seq_len(nrow(f))) {
    sev <- ifelse(is.na(f$severity[[i]]), "Context", as.character(f$severity[[i]]))
    cat(glue("<div class='flag {aesc(sev)}'><strong>{aesc(sev)}: {aesc(f$title[[i]])}</strong><br>{aesc(f$detail[[i]])}</div>"))
  }
}
```

# Peer group

```{r results='asis'}
cat(glue("<div class='peer-box'><strong>{aesc(peer_mode_label)}</strong><br>{aesc(peer_definition)}<br><strong>{peer_count}</strong> facilities are included in the peer benchmark. The target facility itself is excluded from the peer distribution.</div>"))
if (!is.na(peer_count) && peer_count < 10) {
  cat("<div class='peer-warning'><strong>Small peer cohort:</strong> Interpret percentile and quartile comparisons cautiously. A broader cohort may be useful as a sensitivity check.</div>")
}
```

The benchmarking engine supports three peer methods: **automatic matching**, **user-defined filters**, and an **explicit facility list**. This report records the method used so the comparison is reproducible. A future interactive selector can expose these same controls directly to the user without changing the benchmark logic.

```{r}
if (nrow(peer_members)) {
  shown <- peer_members |> slice_head(n = 40)
  peer_gt <- shown |>
    gt() |>
    tab_header(title = md("**Peer facilities used in this benchmark**"), subtitle = if (nrow(peer_members) > 40) paste0("Showing first 40 of ", nrow(peer_members), " peers") else paste0(nrow(peer_members), " peers")) |>
    fmt_number(columns = Beds, decimals = 0) |>
    tab_options(table.font.size = px(11), table.width = pct(100))
  peer_gt
}
```

# Latest benchmark

```{r}
show_metrics <- c(
  "occupancy_rate", "average_daily_census", "medicare_day_share", "medicaid_day_share", "other_day_share",
  "net_patient_revenue_per_day", "operating_expense_per_day", "revenue_expense_spread_per_day",
  "patient_service_margin", "total_margin", "salary_cost_per_day", "contract_labor_per_day",
  "contract_labor_ratio", "total_labor_cost_per_day", "labor_cost_share_of_operating_expense",
  "days_cash_on_hand", "current_ratio", "days_accounts_receivable", "debt_to_assets"
)
bench_table <- b |>
  filter(.data$metric %in% show_metrics) |>
  mutate(
    Facility = mapply(fmt, target_value, format),
    `Peer median` = mapply(fmt, peer_median, format),
    `Peer p25` = mapply(fmt, peer_p25, format),
    `Peer p75` = mapply(fmt, peer_p75, format),
    Difference = mapply(fmt, difference_from_peer_median, format),
    Percentile = percent(target_percentile, accuracy = 1)
  ) |>
  select(Domain = domain, Metric = label, Facility, `Peer median`, `Peer p25`, `Peer p75`, Difference, Percentile)
bench_table |> gt(groupname_col = "Domain") |> tab_options(table.font.size = px(12), table.width = pct(100))
```

# Trends

The blue line and circles are the **target facility**. The orange dashed line and squares are the **peer median**. The gray shaded band is the **peer interquartile range (25th–75th percentile)**. Target points are shown only for years with a usable target cost report.

## Occupancy

```{r fig.height=4.3, fig.width=8.5}
plot_metric("occupancy_rate", "Occupancy rate")
```

```{r results='asis'}
cat(metric_explanation_html("occupancy_rate"))
```

## Medicare day share

```{r fig.height=4.3, fig.width=8.5}
plot_metric("medicare_day_share", "Medicare day share")
```

```{r results='asis'}
cat(metric_explanation_html("medicare_day_share"))
```

## Medicaid day share

```{r fig.height=4.3, fig.width=8.5}
plot_metric("medicaid_day_share", "Medicaid day share")
```

```{r results='asis'}
cat(metric_explanation_html("medicaid_day_share"))
```

## Revenue per resident day

```{r fig.height=4.3, fig.width=8.5}
plot_metric("net_patient_revenue_per_day", "Net patient revenue per resident day")
```

```{r results='asis'}
cat(metric_explanation_html("net_patient_revenue_per_day"))
```

## Operating expense per resident day

```{r fig.height=4.3, fig.width=8.5}
plot_metric("operating_expense_per_day", "Operating expense per resident day")
```

```{r results='asis'}
cat(metric_explanation_html("operating_expense_per_day"))
```

## Patient service margin

```{r fig.height=4.3, fig.width=8.5}
plot_metric("patient_service_margin", "Patient service margin")
```

```{r results='asis'}
cat(metric_explanation_html("patient_service_margin"))
```

## Contract labor ratio

```{r fig.height=4.3, fig.width=8.5}
plot_metric("contract_labor_ratio", "Contract labor ratio")
```

```{r results='asis'}
cat(metric_explanation_html("contract_labor_ratio"))
```

## Total labor cost per resident day

```{r fig.height=4.3, fig.width=8.5}
plot_metric("total_labor_cost_per_day", "Total labor cost per resident day")
```

```{r results='asis'}
cat(metric_explanation_html("total_labor_cost_per_day"))
```

## Days cash on hand

```{r fig.height=4.3, fig.width=8.5}
plot_metric("days_cash_on_hand", "Days cash on hand")
```

```{r results='asis'}
cat(metric_explanation_html("days_cash_on_hand"))
```

# What to validate next

A facility-specific operating review should use these cost-report signals to decide what to investigate next. The highest-value enrichments for the commercial product are PBJ staffing/contract hours, Care Compare quality and survey measures, ownership/chain relationships, SNF VBP/QRP measures, local labor-market data, and reimbursement/referral-market context.

<div class="source-note">
<strong>Source and limitations:</strong> CMS Healthcare Cost Report Information System / Skilled Nursing Facility Cost Report public-use data. Cost reports are screening and benchmarking data, not audited operating recommendations. Facility comparisons depend on reported fields, report-period quality, peer definition, and accounting practices. SNFBenchmarking.com is an independent product and is not endorsed by CMS.
</div>
